% =====================================================================
% Cover Letter — JCIS Submission
% "Contact without Commitment: Atomistic Characterisation of
%  β-Lactoglobulin Adsorption Dynamics at the Air–Water Interface"
% Pornjariyawatch & Pongprayoon, Kasetsart University
%
% Instructions before submitting:
%   1. Insert Zenodo DOI on the \newcommand{\zenododoi} line below
%   2. Update \date{} to submission date
%   3. Compile: pdflatex cover_letter.tex
%   4. Upload cover_letter.pdf to EditorialManager
% =====================================================================

\newcommand{\zenododoi}{10.5281/zenodo.XXXXXX}   % <-- INSERT REAL DOI HERE

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage[margin=1in,top=1.2in,bottom=1.2in]{geometry}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{colorlinks=true,urlcolor=blue!50!black,citecolor=blue!40!black}
\usepackage{parskip}
\setlength{\parskip}{0.8em}
\pagestyle{empty}

\begin{document}

\date{\today}
\noindent\today

\bigskip

\noindent The Editorial Office\\
\textit{Journal of Colloid and Interface Science}\\
Elsevier

\bigskip

Dear Editor,

We submit for your consideration the manuscript \textbf{``Contact without
Commitment: Atomistic Characterisation of $\beta$-Lactoglobulin Adsorption
Dynamics at the Air--Water Interface''} by Chalakon~Pornjariyawatch and
Prapasiri~Pongprayoon (Department of Chemistry, Faculty of Science, Kasetsart
University, Bangkok, Thailand).

$\beta$-Lactoglobulin (BLG), the dominant whey protein in bovine milk, is a
central determinant of dairy foam stability, yet the molecular mechanism by
which it adsorbs at the air--water interface has remained uncharacterised at
atomic resolution. We address this gap with the first unbiased atomistic
molecular dynamics study of native-state BLG at the air--water interface,
accumulating \textbf{4.00~$\mu$s} of trajectory across one bulk-start simulation
and three near-interface replicas using the CHARMM36m force field and a
$12\times12\times35$\,nm slab geometry.

Our principal findings are: (i) BLG contacts the interface frequently ---
\textbf{613 discrete events} across 4.00~$\mu$s, resolved at single-atom
contact distance --- yet 607 of these events terminate without commitment,
revealing a genuine kinetic bottleneck rather than a sampling limitation;
(ii) calyx SASA (24--37~nm$^2$, corrected for periodic-boundary-condition
artefacts) and orientation are \textbf{statistically independent}
($r = +0.006$; block bootstrap 95\% CI $[-0.09, +0.11]$), ruling out simple
two-factor coupling as the commitment trigger; (iii) the protein remains
\textbf{globally compact} throughout 4~$\mu$s of aggregate trajectory,
inconsistent with pre-adsorption global unfolding models; and (iv)
\textbf{Loop CD/EF} (residues 57--60) emerges as the dominant calyx-mobility
region near the interface, replacing Loop BC in the bulk. These results
establish a precise atomistic baseline for the pre-commitment contact ensemble
--- the missing reference state for designing enhanced-sampling calculations
of the adsorption pathway.

We believe this work is well-suited for \textit{JCIS} given the journal's
strong coverage of protein behaviour at fluid interfaces, foam and emulsion
stabilisation mechanisms, and colloid science at the molecular scale. The
findings are directly relevant to the design of functional dairy ingredients
and to the broader question of how globular proteins cross the air--water
interface.

The manuscript has not been submitted elsewhere and is not under consideration
at any other journal. All authors have approved the submission. The authors
declare no competing financial interests.

The dataset supporting this study has been deposited at Zenodo
(DOI:~\href{https://doi.org/\zenododoi}{\zenododoi}), and all analysis scripts
are available at
\href{https://github.com/cpornjar/blg-airwater-interface-md}{github.com/cpornjar/blg-airwater-interface-md}.

We are grateful for your consideration and look forward to your response.

\bigskip

Sincerely,

\bigskip\bigskip

\noindent\textbf{Chalakon Pornjariyawatch}\\
PhD Student, COMFHA Laboratory\\
Department of Chemistry, Faculty of Science\\
Kasetsart University, Bangkok 10900, Thailand\\
\href{mailto:chalakon.p@ku.th}{chalakon.p@ku.th}

\bigskip

\noindent\textbf{Prapasiri Pongprayoon} (corresponding author)\\
Associate Professor, COMFHA Laboratory\\
Department of Chemistry, Faculty of Science\\
Kasetsart University, Bangkok 10900, Thailand

\end{document}
